---
title: "Creating the dataset for predicting stroke occurrence - `r Sys.Date()`"
author: "Sam Henderson-Palmer"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = TRUE, fig.width = 6, fig.align = "center"
)
```

# Raw dataset

The raw data can be found on Kaggle [here](https://www.kaggle.com/datasets/aouatifcherdid/healthcare-dataset-stroke-data) which was downloaded as a CSV and stored locally in the `/Data` folder. 

# Missingness

The raw data was first loaded and checked for missing information:
```{r check_missings}
# Load packages
library(tidyverse)

# Read in raw data
stroke_raw <- read.csv("healthcare-dataset-stroke-data.csv")

# Get overview of data structure, numerical ranges and categories
str(stroke_raw)
summary(stroke_raw)
table(stroke_raw$stroke)
table(stroke_raw$gender)
table(stroke_raw$ever_married)
table(stroke_raw$work_type)
table(stroke_raw$Residence_type)
table(stroke_raw$smoking_status)

# Check for NAs
which(is.na(stroke_raw))
```
# Cleaning

Then some data cleaning ahead of modelling, for example changing categorical variables to factors and ensuring our target variable `stroke` is a factor with the first level as event of interest i.e. stroke:
```{r cleaning}
stroke_raw %>% 
  # Convert 1's and 0's for stroke for clarity 
  mutate(stroke = replace(stroke, stroke == 1, "Stroke"),
         stroke = replace(stroke, stroke == 0, "No stroke")) %>% 
  # Convert variable types to categorical
  mutate(across(c(gender, hypertension, heart_disease, ever_married, 
                  work_type, Residence_type, smoking_status, stroke), as.factor)) -> stroke

# Reverse order of stroke factor levels so first level is event of interest i.e. stroke
stroke$stroke <- factor(stroke$stroke, levels = rev(levels(stroke$stroke)))

# Check summary of cleaned data
summary(stroke)
str(stroke)
```

# Saving dataset
Once we have the final dataset prepared we just need to save it into RDS format which preserves our the datatypes and metadata:
```{r save_data}
# Save cleaned dataset
saveRDS(stroke, file = "stroke_complete.RDS")
```

